---
title: "Hands-on Exercise 3"
author: "Jenna Cheo"
---

## Getting Started

In this exercise, we will use the following datasets:

-   `CHILDCARE`, a point feature data providing both location and attribute information of childcare centres. It was downloaded from Data.gov.sg and is in geojson format.

-   `MP14_SUBZONE_WEB_PL`, a polygon feature data providing information of URA 2014 Master Plan Planning Subzone boundary data. It is in ESRI shapefile format. This data set was also downloaded from data.gov.sg.

-   `CostalOutline`, a polygon feature data showing the national boundary of Singapore. It is provided by SLA and is in ESRI shapefile format.

### Installing and Loading R packages

```{r}
pacman::p_load(sf, raster, spatstat, tmap, tidyverse)
```

```{r}
mpsz_sf <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

### Importing Datasets

```{r}
childcare_sf <- st_read("data/aspatial/child-care-services-geojson.geojson") %>% st_transform(crs =3414)
```

```{r}
mpsz_sf <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
sg_sf <- st_read(dsn = "data/geospatial", layer="CostalOutline")
```

## Geospatial Data Wrangling

DIY: Using the method you learned in Lesson 2, assign the correct crs to mpsz_sf and sg_sf simple feature data frames.

```{r}
childcare_sf <- st_transform(childcare_sf, crs = 3414)
mpsz_sf <- st_transform(mpsz_sf, crs = 3414)
```

### Mapping Geospatial Datasets

DIY: Using the mapping methods you learned in Hands-on Exercise 3, prepare a map as shown below.

```{r}
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()
```

```{r}
tmap_mode('plot')
```

Notice that at the interactive mode, **tmap** is using [**leaflet for R**](https://rstudio.github.io/leaflet/) API. The advantage of this interactive pin map is it allows us to navigate and zoom around the map freely. We can also query the information of each simple feature (i.e. the point) by clicking of them. Last but not least, you can also change the background of the internet map layer. Currently, three internet map layers are provided. They are: ESRI.WorldGrayCanvas, OpenStreetMap, and ESRI.WorldTopoMap. The default is ESRI.WorldGrayCanvas.

> Reminder: Always remember to switch back to plot mode after the interactive map. This is because, each interactive mode will consume a connection. You should also avoid displaying ecessive numbers of interactive maps (i.e. not more than 10) in one RMarkdown document when publish on Netlify.

### Converting sf data frames to sp’s Spatial\* class
