---
title: "Hands on ex 12"
author: "Jenna Cheo"
date: "November 1, 2024"
date-modified: "last-modified"
format: html
editor: visual
---

# Geographically Weighted Predictive Models

## Getting Started

```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse)
```

```{r}
mdata <- read_rds("data/rds/mdata.rds")
```

## Data Wrangling

### Data Sampling

```{r}
set.seed(1234)
resale_split <- initial_split(mdata, 
                              prop = 6.5/10,)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

```{r}
write_rds(train_data, "data/rds/train_data.rds")
write_rds(test_data, "data/rds/test_data.rds")
```

## Computing Correlation Matrix

```{r}
mdata_nogeo <- mdata %>%
  st_drop_geometry()
corrplot::corrplot(cor(mdata_nogeo[, 2:17]), 
                   diag = FALSE, 
                   order = "AOE",
                   tl.pos = "td", 
                   tl.cex = 0.5, 
                   method = "number", 
                   type = "upper")
```

The correlation matrix above shows that all the correlation values are below 0.8. Hence, there is no sign of multicolinearity.

## Retrieving stored data

```{r}
train_data <- read_rds("data/rds/train_data.rds")
test_data <- read_rds("data/rds/test_data.rds")
```

## Building a non-spatial multiple linear regression

```{r}
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
summary(price_mlr)
```

```{r}
write_rds(price_mlr, "data/rds/price_mlr.rds" ) 
```

## Preparing coordinates data

### Extracting coordinate data

```{r}
coords <- st_coordinates(mdata)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```

```{r}
coords_train <- write_rds(coords_train, "data/rds/coords_train.rds" )
coords_test <- write_rds(coords_test, "data/rds/coords_test.rds" )
```

### Dropping geometry field

```{r}
train_data <- train_data %>% 
  st_drop_geometry()
```

## Calibrating Random Forest Model

```{r}
set.seed(1234)

rf <- ranger(resale_price ~ floor_area_sqm + storey_order + 
               remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE + 
               PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL + 
               PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
               WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
               WITHIN_1KM_PRISCH,
             data=train_data)
rf
```

```{r}
write_rds(rf, "data/rds/rf.rds")
```

```{r}
rf <- read_rds("data/rds/rf.rds")
rf
```

## Calibrating Geographical Random Forest Model

### Calibrating using training data

```{r}
set.seed(1234)
gwRF_adaptive <- grf(formula = resale_price ~ floor_area_sqm + storey_order +
                       remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE +
                       PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL +
                       PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                       WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                       WITHIN_1KM_PRISCH,
                     dframe=train_data, 
                     bw=55,
                     kernel="adaptive",
                     coords=coords_train)
```

```{r}
write_rds(gwRF_adaptive, "data/rds/gwRF_adaptive.rds")
```

```{r}
gwRF_adaptive <- read_rds("data/rds/gwRF_adaptive.rds")
```

### Predicting by using test data

unfortunately, im unable to evaluate the last few chunks of code as it takes too long to run and fails.

```{r}
#| eval: false

test_data <- cbind(test_data, coords_test) %>%
  st_drop_geometry()
```

```{r}
#| eval: false

gwRF_pred <- predict.grf(gwRF_adaptive, 
                           test_data, 
                           x.var.name="X",
                           y.var.name="Y", 
                           local.w=1,
                           global.w=0)
```

```{r}
#| eval: false

GRF_pred <- write_rds(gwRF_pred, "data/rds/GRF_pred.rds")
```

Convert it to data frame...

```{r}
#| eval: false

GRF_pred <- read_rds("data/rds/GRF_pred.rds")
GRF_pred_df <- as.data.frame(GRF_pred)
```

```{r}
#| eval: false

test_data_p <- cbind(test_data, GRF_pred_df)
```

```{r}
#| eval: false

write_rds(test_data_p, "data/rds/test_data_p.rds")
```

Calculating Root Mean Square Error (RMSE)

```{r}
#| eval: false

rmse(test_data_p$resale_price, 
     test_data_p$GRF_pred)
```

```{r}
#| eval: false

ggplot(data = test_data_p,
       aes(x = GRF_pred,
           y = resale_price)) +
  geom_point()
```
