---
title: "In Class Exercise 2: IS415-GAA"
author: "Jenna Cheo"
date: "August 26, 2024"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  freeze: true
---

## Getting started

The code chunk below will be used to install and load these packages in RStudio.

```{r}
pacman::p_load(sf, tidyverse)
```

### Importing the data sets into R

One data set will be used in this in class exercise.

-   GEOSPATIAL: Master Plan 2014 Subzone Boundary (Web) (i.e. `MP14_SUBZONE_WEB_PL`) in SHP and KML format \[ [data.gov.sg](https://data.gov.sg/) \]

The code chunk below imports shapefile.

```{r}
mpsz14_shp <- st_read(dsn = "data/MasterPlan2014SubzoneBoundaryWebSHP", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
class(mpsz14_shp)
```

The original KML file provided by data.gov.sg was corrupted and thus was unable to be read using sf.

To resolve this, we rewrite over the file. We also rename the file to a more readable file name.

```{r}
#| output: false
st_write(mpsz14_shp,
         "data/MP14_SUBZONE_WEB_PL.kml",
          delete_dsn = TRUE)
```

The code chunk below imports KML file into R.

```{r}
#| eval: false
mpsz14_kml <- st_read("data/MP14_SUBZONE_WEB_PL.kml")
```
