---
title: "In-class_Ex04"
author: "Jenna Cheo"
date: "9th September 2024"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  freeze: true
  eval: false
---

## Getting Started

### Importing packages

For the purpose of this study, 6 R packages will be used.

They are sf, raster, spatstat, sparr, tmap, tidyverse.

```{r}
pacman::p_load(sf, st, tidyverse, raster, tmap, spatstat, sparr, dplyr)
```

### Importing Dataset

The file is in ESRI shapefile format.

```{r}
#| eval: false
kbb <- st_read(dsn="data/rawdata",
               layer="Kepulauan_Bangka_Belitung")
```

Next, we import the forest fire data which is in CSV format.

```{r}
fire_sf <- read_csv("data/rawdata/forestfires.csv") %>%
  st_as_sf(coords =c("longitude", "latitude"),
           crs = 4326) %>%
st_transform(crs = 32748)
```

```{r}
fire_sf <- fire_sf %>%
  mutate(DayofYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date,
                           label = TRUE,
                           abbr = FALSE))
```

## Data Wrangling

if there are z entities, you most likely will have to drop it (it will cause issues to this exercise in this case)

The revised code chunk is as shown below.

```{r}
kbb_sf <- st_read(dsn="data/rawdata",
               layer="Kepulauan_Bangka_Belitung") %>%
st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748)
```

### Converting OWIN

```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```

`class()` is used to confirm if the output is indeed an owin object

```{r}
class(kbb_owin)
```

## Overall Plot

always polygon first, followed by dot then line

```{r}
tm_shape(kbb_sf) +
  tm_polygons() +
tm_shape(fire_sf) +
  tm_dots()
```

```{r}
tm_shape(kbb_sf) +
  tm_polygons() +
tm_shape(fire_sf) +
  tm_dots(size = 0.1) +
tm_facets(by = "Month_fac",
          free.coords = FALSE,
          drop.units = TRUE)
```

## Computing STKDE by month

```{r}
fire_month <- fire_sf %>%
  dplyr::select(Month_num)
```

```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```

```{r}
summary(fire_month_ppp)
```

```{r}
any(duplicated(fire_month_ppp))
```

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```

### Computing Spatio-Temporal KDE

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

```{r}
tims <- c(7, 8, 9, 10, 11, 12)
par(mfcol=c(2,3))
for (i in tims){
  plot(st_kde, i,
       override.par = FALSE,
       fix.range = TRUE,
       main = paste("KDE at month", i))
}
```

fire_yday_ppp \<- fire_sf %\>% select(DayofYear)

kde_yday \<- spattemp.density( fire_yday_owin) summary(kde_yday)

plot(kde_yday)
