---
title: "Take-home_Ex01"
author: "Jenna Cheo"
date: "5th September 2024"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
---

## 1.0 Introduction

Millions of people have their lives shattered by armed conflict – wars – every year.

Armed conflict has been on the rise since about 2012, after a decline in the 1990s and early 2000s. First came conflicts in Libya, Syria and Yemen, triggered by the 2011 Arab uprisings. Libya’s instability spilled south, helping set off a protracted crisis in the Sahel region. A fresh wave of major combat followed: the 2020 Azerbaijani-Armenian war over the Nagorno-Karabakh enclave, horrific fighting in Ethiopia’s northern Tigray region that began weeks later, the conflict prompted by the Myanmar army’s 2021 power grab and Russia’s 2022 assault on Ukraine. Add to those 2023’s devastation in Sudan and Gaza. Around the globe, more people are dying in fighting, being forced from their homes or in need of life-saving aid than in decades.

Source: [10 Conflicts to Watch in 2024](https://www.crisisgroup.org/global/10-conflicts-watch-2024)

## 2.0 Importing Packages

Before we start the exercise, we will need to import necessary R packages first. We will use the following packages:

-   [`sf`](https://rdrr.io/github/r-spatial/sf/man/sf-package.html) : provides a standardised way to encode spatial vector data in R environment, facilitating spatial data operations and analysis.

-   [`st`](https://rdrr.io/cran/sf/man/st.html) : creats simple features from numeric vectors, matrices, or lists, enabling the representation and manipulation of spatial structures in R.

-   [`spatstat`](https://cran.r-project.org/web/packages/spatstat/): A package for statistical analysis of spatial data, specifically Spatial Point Pattern Analysis. This package was provided by Baddeley, Turner and Ruback (2015) and gives a comprehensive list of functions to perform 1st- and 2nd-order spatial point patterns analysis and derive kernel density estimation (KDE) layer.

-   [`tidyverse`](https://www.tidyverse.org/) : a collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structure.

-   [`raster`](https://cran.r-project.org/web/packages/raster/) : reads, writes, manipulates, analyses and model of gridded spatial data (i.e. raster). In this hands-on exercise, it will be used to convert image output generate by spatstat into raster format.

-   [`tmap`](https://cran.r-project.org/web/packages/tmap/): Packages used for creating static and interactive visualisations summary statistics and KDE layers.

-   [`spatstat`](https://spatstat.org/) : for spatial statistics with a strong focus on analysing spatial point patterns.

```{r}
pacman::p_load(sf, st, tidyverse, raster, tmap, ggplot2, spatstat)
```

## 3.0 Importing Dataset into R environment

### 3.1 Datasets

For the purpose of this assignment, **armed conflict data** of Myanmar between 2021-2024 from [Armed Conflict Location & Event Data (ACLED)](https://acleddata.com/), an independent, impartial, international non-profit organization collecting data on violent conflict and protest in all countries and territories in the world, should be used.

In terms of event types, students should focus on at least four main event types, namely: Battles, Explosion/Remote violence, Strategic developments, and Violence against civilians.

In terms of study period, students should focus on quarterly armed conflict events from January 2021 until June 2024.

### 3.2 Importing the Myanmar Armed Conflict Dataset

The `ACLED_Myanmar` dataset, which we downloaded from ACLED, is in csv format. To use this data in an R-environment, we need to import it as an sf object. We can do this using the **`st_read()`** function of the **`sf`** package. This function reads the csv data and returns an **`sf`** object that can be used for further analysis.

```{r}
acled_sf <- st_read("data/ACLED_Myanmar.csv") %>%
  st_as_sf(coords = c(
    "longitude", "latitude"),
    crs=4326) %>%
  st_transform(crs = 32647) %>%
  mutate(event_date = dmy(event_date))
```
