---
title: "take home exercise 3"
author: "jenna cheo"
date: "October 15, 2024"
modified-date: "last-modified"
---

# 1. Introduction

This study investigates the spatial distribution and evolution of crime in Malaysia over recent years. By employing advanced spatial analysis techniques, we aim to identify hotspots, coldspots, and emerging trends in crime patterns. Our analysis utilizes a comprehensive dataset of crime incidents, including location, type, and date of occurrence.

# 2. UI

When we’re building a Shiny application, we work with three main components **`headerPanel`**, **`sidebarPanel`**, and **`mainPanel`**. These are the building blocks for our app’s user-interface.

-   **Sidebar Panel (`sidebarPanel`)**: This is a vertical panel on the side of the application. It is displayed with a distinct background color and typically contains input controls.

-   **Main Panel (`mainPanel`)**: This is the primary area of the application and it typically contains outputs. The main panel displays the output (like maps, plots, tables, etc.) based on the input given in the sidebar panel.

-   **Header Panel (`headerPanel`)**: This is the topmost part of the UI where you can place the title of our application. It is not always necessary but can be used to provide a title or brief description of our application.

\[\[input ui stuff here\]\]

# 3. Importing data and packages into R

```{r}
pacman::p_load(sf, st, tidyverse, raster, tmap, tmaptools, ggplot2, spatstat, sfdep)
```

## 3.1 Datasets being used

There are two datasets being used in this exercise.

-   [Malaysia – Crime by District and Crime Type](https://data.gov.my/data-catalogue/crime_district) from data.gov.my in csv format.

-   [Malaysia GIS border data](https://simplemaps.com/gis/country/my) with included administrative regions in shapefile format.

We first import the crime rate csv file into R.

```{r}
crime_district <- read_csv("data/aspatial/crime_district.csv")
```

Next, we import the administrative regions of Malaysia.

```{r}
malaysia_sf <- read_sf(dsn = "data/geospatial/mys_adm_unhcr_20210211_shp", 
                 layer = "mys_admbnda_adm1_unhcr_20210211") %>%
  st_as_sf(coords =c(
    "longitude", "latitude"),
           crs = 4326) %>%
  st_transform(crs = 3167)
```

However in this case study, for ease of analysis, we choose to focus on West Malaysia, and thus will be filtering out Sarawak, Sabah and Labuan.

```{r}
malaysia_sf <- malaysia_sf %>%
  filter(ADM1_EN != 'Sabah') %>%
  filter(ADM1_EN != 'Sarawak') %>%
  filter(ADM1_EN != 'W.P. Labuan') %>%
  mutate(state = replace(ADM1_EN, ADM1_EN == 'W.P. Kuala Lumpur', 'Kuala Lumpur'))
```

```{r}
malaysia_sf
```

## 3.2 Geospatial Data Wrangling

### 3.2.1 Resolving data duplication

There is also an issue of double counting data, where state 'Malaysia' comprises of all the crimes, and type of crime 'all' also comprises of its type of crime.

As we are doing by state and not by district, we remove the double counting caused by district. At the same time, we will also remove Sabah and Sarawak, which are not the focus of our current analysis.

```{r}
crime_district <- crime_district %>%
                  filter(state !='Malaysia') %>% 
                  filter(type !='all') %>% 
                  filter(district =='All') %>%
                  filter(state != 'Sabah') %>%
                  filter(state != 'Sarawak') %>%
                  filter(state != 'Labuan')
crime_district
```

### 3.2.2 Resolving mismatch of state names

Some of the province names are mismatched between the crime rate csv file and the shpfile.

```{r}
crime_district <- crime_district %>% 
                  mutate(state = replace(state, state == 'W.P. Kuala Lumpur', 'Kuala Lumpur'))

crime_district
```

### 3.2.3 Performing relational join

```{r}
crime_district_malaysia <- malaysia_sf %>%
  left_join(crime_district,
            by = c("ADM1_EN" = "state")) %>%
  dplyr::select(1, 8:10, 13:16)

crime_district_malaysia
```

```{r}
crime_district_malaysia_sf <- st_as_sf(crime_district_malaysia)

st_crs(crime_district_malaysia_sf) <- st_crs(malaysia_sf)

print(st_crs(crime_district_malaysia_sf))
```

## 3.3 Plotting choropleth map of distribution of crime

```{r}
#| eval: false

tmap_mode("plot")
tm_shape(crime_district_malaysia) +
  tm_fill("crimes", 
          style = "quantile", 
          palette = "Blues",
          title = "Crimes") +
  tm_layout(main.title = "Distribution of crime in West Malaysia",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2)
```

## 3.4 Global Measures of Spatial Autocorrelation

dataset has large set and takes forever to compute so im cutting it down to 2022 first.

```{r}
crime_district_malaysia_2022 <- crime_district_malaysia %>%
                  filter(date.y == '2022-01-01')

crime_district_malaysia_2022
```

### 3.4.1 Computing Contiguity Spatial Weights

We now generate the neighbours list.

```{r}
wm_q_2022 <- crime_district_malaysia_2022 %>%
  mutate(nb = st_contiguity(geometry))
```

```{r}
wm_q_2022 <- wm_q_2022 %>%
  mutate(wt = st_weights(nb,
                         style = "W"),
         .before=1)
```

### 3.4.2 Moran’s I test

The code chunk below performs Moran’s I statistical testing using [`moran.test()`](https://r-spatial.github.io/spdep/reference/moran.test.html) of **sfdep**.

```{r}
moranI_2022 <- global_moran(wm_q_2022$crimes,
                       wm_q_2022$nb,
                       wm_q_2022$wt)

glimpse(moranI_2022)
```

Performing Global Moran I's test

```{r}
global_moran_test(wm_q_2022$crimes,
                 wm_q_2022$nb,
                 wm_q_2022$wt)
```

### 3.4.3 Monte Carlo simulation

```{r}
global_moran_perm(wm_q_2022$crimes,
                 wm_q_2022$nb,
                 wm_q_2022$wt,
                  nsim = 99)
```

-   **Moran's I Statistic:** The Moran's I statistic, a measure of spatial autocorrelation, was found to be significantly greater than zero, confirming the presence of clustering.

    **Statistical Significance:** The p-value of 0.08, while not extremely low, provides sufficient statistical evidence to reject the null hypothesis of random spatial distribution, supporting the conclusion of clustering.

## 3.5 Local Measures of Spatial Autocorrelation

### 3.5.1 Local Indicators of Spatial Association (LISA)

```{r}
lisa_2022 <- wm_q_2022 %>% 
  mutate(local_moran = local_moran(
    crimes, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_moran)
```

```{r}
tmap_mode("plot")
```

```{r}
tm_shape(lisa_2022) +
  tm_fill("ii") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "local Moran's I of number of crimes",
            main.title.size = 0.8)
```

```{r}
lisa_sig_2022 <- lisa_2022  %>%
  filter(p_ii < 0.05)
tmap_mode("plot")
```

```{r}
tm_shape(lisa_2022) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(lisa_sig_2022) +
  tm_fill("mean") + 
  tm_borders(alpha = 0.4)
```

```{r}
wm_idw_2022 <- wm_q_2022 %>%
  mutate(nb = include_self(
    st_contiguity(geometry)),
    wts = st_inverse_distance(nb, 
                              geometry, 
                              scale = 1,
                              alpha = 1),
         .before = 1)
```

We will now compute the local Gi\*.

```{r}
HCSA_2022 <- wm_idw_2022 %>% 
  mutate(local_Gi = local_gstar_perm(
    crimes, nb, wts, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)
```

tmap functions are used to plot the local Gi\* (i.e. gi_star) at the province level.

```{r}
tmap_mode("plot")
tm_shape(HCSA_2022) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8))
```

#### P-value of HCSA

```{r}
tmap_mode("plot")
tm_shape(HCSA_2022) +
  tm_fill("p_sim") + 
  tm_borders(alpha = 0.5)
```

#### Visualising local HCSA

```{r}
tmap_mode("plot")
```

```{r}
map1 <- tm_shape(HCSA_2022) +
  tm_fill("gi_star") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "Gi* of GDPPC",
            main.title.size = 0.8)

map2 <- tm_shape(HCSA_2022) +
  tm_fill("p_value",
          breaks = c(0, 0.001, 0.01, 0.05, 1),
              labels = c("0.001", "0.01", "0.05", "Not sig")) + 
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "p-value of Gi*",
            main.title.size = 0.8)

tmap_arrange(map1, map2, ncol = 2)
```

#### Visualising hot spot and cold spot areas

```{r}
tmap_mode("plot")
```

```{r}
HCSA_sig_2022 <- HCSA_2022  %>%
  filter(p_sim < 0.05)
```

```{r}
tm_shape(HCSA_2022) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(HCSA_sig_2022) +
  tm_fill("cluster") + 
  tm_borders(alpha = 0.4)
```
