---
title: "Hands On Exercise 2"
author: "Jenna Cheo"
date: "26th September 2024"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  freeze: true
---

# 1.0 Introduction

Drug abuse is associated with significant negative health, financial and social consequences. Yet, illicit drug consumption remains highly prevalent and continues to be a growing problem worldwide.

The geopolitics of Thailand which is near the [Golden Triangle](https://en.wikipedia.org/wiki/Golden_Triangle_(Southeast_Asia)) of Indochina, the largest drug production site in Asia, and the constant transportation infrastructure development made Thailand became market and transit routes for drug trafficking to the third countries.

In Thailand, drug abuse is one of the major social issue. There are about 2.7 million youths using drugs in Thailand. Among youths aged between 15 and 19 years, there are about 300,000 who have needs for drug treatment. Most of Thai youths involved with drugs are vocational-school students, which nearly doubles in number compared to secondary-school students.

# 2.0 Getting Started

## 2.1 Loading packages into R

```{r}
pacman::p_load(sf, st, tidyverse, raster, tmap, tmaptools, ggplot2, spatstat, sfdep)
```

## 2.2 Importing the datasets

For the purpose of this take-home exercise, two data sets shall be used, they are:

-   [Thailand Drug Offenses \[2017-2022\]](https://www.kaggle.com/datasets/thaweewatboy/thailand-drug-offenses-2017-2022) at Kaggle.

-   [Thailand - Subnational Administrative Boundaries](https://data.humdata.org/dataset/cod-ab-tha?) at HDX. You are required to use the province boundary data set.

### 2.2.1 Importing Thailand Drug Offenses (csv file)

The dataset, which we downloaded from Kaggle, is in csv format. The codes chunk below uses `read_csv()` function of **readr** package to import *the dataset* into R as a tibble data frame called *drug_offences*.

The data has also been transformed such it only contains cases of drug use.

```{r}
drug_offences <- read_csv("data/thai_drug_offenses_2017_2022.csv")  %>%
                  filter(types_of_drug_offenses == 'drug_use_cases')
```

```{r}
glimpse(drug_offences)
```

### 2.2.2 Importing Thailand Subnational Administrative Boundaries (shp)

```{r}
thailand_sf <- read_sf(dsn = "data", 
                 layer = "tha_admbnda_adm1_rtsd_20220121") %>%
  st_as_sf(coords =c(
    "longitude", "latitude"),
           crs = 4326) %>%
  st_transform(crs = 24047)
```

```{r}
thailand_sf
```

## 2.3 Geospatial Data Wrangling

### 2.3.1 Fixing province names

There is a mismatch of the province names between the datasets we are importing. We resolve this by changing the names of provinces in the drug offenses dataset to align.

-   Loburi -\> Lop Buri

-   Buogkan -\> Bueng Kan

```{r}
drug_offences <- drug_offences %>% 
                  mutate(province_en = replace(province_en, province_en == 'buogkan', 'Bueng Kan')) %>%
                  mutate(province_en = replace(province_en, province_en == 'Loburi', 'Lop Buri'))
drug_offences
```

### 2.3.2 Performing relational join

The code chunk below will be used to update the attribute table of *thailand_sf*’s SpatialPolygonsDataFrame with the attribute fields of *drug_offenses* dataframe. This is performed by using *left_join()* of **dplyr** package. We have filtered it out to contain

```{r}
drug_offences_thailand <- thailand_sf %>%
  left_join(drug_offences,
            by = c("ADM1_EN" = "province_en",
                  "ADM1_TH" = "province_th")) %>%
  dplyr::select(1:2, 17:20)
```

```{r}
drug_offences_thailand
```

```{r}
class(drug_offences_thailand)
```

### 2.3.3 Visualising Regional Development Indicator

We prepare a basemap and a choropleth map showing the distribution of drug offences by using `qtm()` of **tmap** package.

```{r}
equal <- tm_shape(drug_offences_thailand) +
  tm_fill("no_cases",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal interval classification")

quantile <- tm_shape(drug_offences_thailand) +
  tm_fill("no_cases",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal quantile classification")

tmap_arrange(equal, 
             quantile, 
             asp=1, 
             ncol=2)
```

# 3.0 Global Measures of Spatial Autocorrelation

### 3.0.1 Computing Contiguity Spatial Weights

We need to first construct a spatial weight of the study area, used to define the neighbourhood relationships between the provinces in Thailand.

```{r}
wm_q <- drug_offences_thailand %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb,
                         style = "W"),
         .before=1)
```

## 3.2 Moran’s I test

The code chunk below performs Moran’s I statistical testing using [`moran.test()`](https://r-spatial.github.io/spdep/reference/moran.test.html) of **sfdep**.

```{r}
moranI <- drug_offences_thailand(wm_q$no_cases,
                       wm_q$nb,
                       wm_q$wt)

glimpse(moranI)
```

Performing Global Moran’s I test

```{r}
global_moran_test(wm_q$no_cases,
                  wm_q$nb,
                  wm_q$wt)
```

### 3.2.1 Monte Carlo simulation

Next, `global_moran_perm()` is used to perform Monte Carlo simulation.

```{r}
setseed(1234)

global_moran_perm(wm_q$no_cases,
                       wm_q$nb,
                       wm_q$wt,
                  nsim = 99)
```

The statistical report on previous tab shows that the p-value is smaller than alpha value of 0.05. Hence, we have enough statistical evidence to reject the null hypothesis that the spatial distribution of drug cases are resemble random distribution (i.e. independent from spatial).

Because the Moran's I statistics is greater than 0. We can infer that the spatial distribution shows sign of clustering.

# 4.0 Local Measures of Spatial Autocorrelation

## 4.1 Local Indicators of Spatial Association (LISA)

```{r}
lisa <- wm_q %>% 
  mutate(local_moran = local_moran(
    no_cases, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_moran)
```

```{r}
tmap_mode("plot")
```

The output of `local_moran()` is a sf dataframe containing the following:

-   *ii*: local moran statistic
-   *eii*: expectation of local moran statistic
-   *var_ii*: variance of local moran statistic
-   *z_ii*: standard deviate of local moran statistic
-   *skewness*: the output of e1071::skewness() for the permutation samples underlying the standard deviates
-   *kurtosis*: For `localmoran_perm`, the output of e1071::kurtosis() for the permutation samples underlying the standard deviates.

In this code chunk below, tmap functions are used prepare a choropleth map by using value in the *ii* field.

```{r}
tm_shape(lisa) +
  tm_fill("ii") + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "local Moran's I of number of drug cases",
            main.title.size = 0.8)
```

In lisa sf data.frame, we can find three fields contain the LISA categories. They are *mean*, *median* and *pysal*. In general, classification in *mean* will be used as shown in the code chunk below.

```{r}
lisa_sig <- lisa  %>%
  filter(p_ii < 0.05)
tmap_mode("plot")
```

```{r}
tm_shape(lisa) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(lisa_sig) +
  tm_fill("mean") + 
  tm_borders(alpha = 0.4)
```

we will need to derive a spatial weight matrix before we can compute local Gi\* statistics

```{r}
wm_idw <- hunan_GDPPC %>%
  mutate(nb = include_self(
    st_contiguity(geometry)),
    wts = st_inverse_distance(nb, 
                              geometry, 
                              scale = 1,
                              alpha = 1),
         .before = 1)
```
